### 2020-06-24 manuscript/manuscript.rmd
rewrote code for figure 3, some text edits

## Variation in models and sampling across published studies

@@

High-throughput sequencing allows investigators to readilly assemble matrices with hundreds or thousands of protein-coding genes (Extended Data Table 1). Studies of animal phylogeny have used a wide variety of different approaches to identifying and selecting genes for their matrices. These take different approaches to sequence comparison, filtering, and orthology assessment. As a result, the genes selected for analysis varies widely between studies (Fig. 2, Extended Data Fig. 1). To explore this variation in gene sampling, we annotated each gene with against SwissProt [@boeckmann:2003swiss] and Gene Ontology databases (Supplementary Table XX). These annotations show XXX. We identified the number of ribosomal protein genes in each data matrix (Extended Data Fig. 2A). We also identified BUSCO genes [@waterhouse2017busco] in each matrix (Extended Data Fig. 2B). The animal BUSCO dataset contains 978 genes that tend to occur as a single copy used to assess the completeness of genome assemblies [@waterhouse2017busco]. All matrices have than 50% of the BUSCO genes, and several less than 20%. The fact that so many BUSCO genes are not in these matrices reflects that the gene selection criteria include many additional factors, such as high expression (since so many phylogenetomic data are based on transcriptome data) and features of molecular evolution.

```{r figure_taxon_rectangles, fig.height = 18, fig.width = 20 }

@@

## Variation in published results

```{r figure_support_analyses_published}
analyses_published=read_tsv("../data_processed/tables/previously_published_analyses.tsv")
analyses_published_filter = dplyr::filter(analyses_published, !grepl('Recoding', model_combined))

x1 = factor(analyses_published_filter$model_combined, levels=c("WAG", "LG", "GTR", "data partitioning", "Recoding + GTR", "Recoding + GTR + CAT", "Poisson + CAT", "GTR + CAT"))
p1 = analyses_published_filter %>%
	ggplot( aes(x=x1, y=clade, col=final, shape=inference) ) +
		geom_jitter( width = 0.20, height = 0.20, alpha=0.7, size=2 ) +
		scale_colour_manual(values = result_colors) +
		theme_classic() + theme(axis.text.x = element_text(angle=45, hjust=1))+
   xlab("Model complexity") + ylab("Clade")

analyses_new=analyses_new %>%
mutate(model=recode(model,
                    `GTR20`="GTR",
                    `CAT+F81`="Poisson+CAT"))

x2 = factor(analyses_new$model, levels=c("WAG", "GTR", "Poisson+C60", "WAG+C60", "LG+C60", "Poisson+CAT", "GTR+CAT"))
p2 = analyses_new %>%
	ggplot( aes(x=x2, y=clade, col=result, shape=inference) ) +
		geom_jitter( width = 0.20, height = 0.20, alpha=0.7, size=2 ) +
		scale_colour_manual(values = result_colors) +
		theme_classic() + theme(axis.text.x = element_text(angle=45, hjust=1))+
       xlab("Model complexity") + ylab("Clade")

ggarrange(p1, p2, ncol = 2, labels = c("A", "B"), common.legend = TRUE, legend = "right")

```
**Fig. 3.**  A summary of phylogenomic results from the previous studies and reanalysis conducted in this study. Each shape represents an phylogenetic analysis that has been analyzed by ML (designated by the circle) or BI inference (designated by the triangle). Analyses with a support of bootstrap values less than 90% or a posterior probability less than 0.95 are considered as unresolved. (A) A total of `r nrow(analyses_published_filter)` analyses were transcribed from the literature (related to Supplementary Table XX). (B) A total of `r nrow(analyses_new)` new phylogenomic analyses were conducted in this study (related to Supplementary Table XX).

Here we summarized all the `r nrow(analyses_published)` previous phylogenetic analyses from `r length(unique(analyses_published$manuscript))` studies (Fig. 3A and Supplementary Table 2). Among these studies, the main conclusions of five studies strongly in favor of Porifera-sister and 10 in support of Ctenophora-sister (Extended Data Table 1). Moreover, three studies are secondary studies without any new data matrices. Interestingly, we found that all analyses strongly support of Ctenophora-sister once site-homogeneous models were used, no matter what outgroup choice, phylogenetic inference and whether data-recoding are used. The support for Porifera-sister increases with the exclusion of more distantly related outgroups and with the use of site-heterogeneious CAT (Poisson+CAT or GTR+CAT) models in several matrices, including Philippe2009, Ryan2013_est and Whelan2015. It should be noted that many BI analyses with CAT model are not well resolved (either not reach convergence or do not provide significant support for the key node of animal evolution) mainly due to computational limits of CAT models employed in PhyloBayes. The strong support of Porifera-sister can only be recovered from analyses using Poisson+CAT and recoding methods with GTR+CAT models in several matrices. Although GTR+CAT model generally has better model adequacy than Poisson+CAT or site-homogeneous models from previous studies[@pisani2015genomic; @feuda2017improved], Porifera-sister is not always supported since several analyses strongly in favor Ctenophora-sister even with only choanofalgellates are used as outgroups (Fig. 3A). Thus, our summary of previous phylogenetic analyses strongly indicated that the trend of the support of Porifera-sister is generally increasing with the use of more complex CAT models and more closely related outgroups in several representative matrices, whereas Ctenophora-sister is supported in most other conditions.

@@

``` {r sensitive, fig.width=8, fig.height=8}

>>>>

## Variation in models and sampling across published analyses

@@

High-throughput sequencing allows investigators to readilly assemble matrices with hundreds or thousands of protein-coding genes (Extended Data Table 1). Studies of animal phylogeny have used a wide variety of different approaches to identifying and selecting genes for their matrices. As a result, the genes selected for analysis vary widely (Fig. 2, Extended Data Fig. 1). To explore this variation in gene sampling, we annotated each gene with against SwissProt [@boeckmann:2003swiss] and Gene Ontology databases (Supplementary Table XX). These annotations show XXX. We identified the ribosomal protein and BUSCO genes [@waterhouse2017busco] in each data matrix (Extended Data Fig. 2). The animal BUSCO dataset contains 978 genes that tend to occur as a single copy used to assess the completeness of genome assemblies [@waterhouse2017busco]. All matrices have than 50% of the BUSCO genes, and several less than 20%. The fact that so many BUSCO genes are not in these matrices reflects that the gene selection criteria include many additional factors, such as high expression (since so many phylogenetomic data are based on transcriptome data) and features of molecular evolution.

```{r figure_taxon_rectangles, fig.height = 7.2, fig.width = 7.2 }

@@

## Variation in results across published analyses

```{r figure_support_analyses, fig.width=7.2, fig.height=3.6}

analyses_published_processed = 
  analyses_published %>%
  dplyr::filter( !grepl( 'Recoding', model_combined ) ) %>%
  dplyr::rename( model = model_combined ) %>%
  select( model, clade, final, inference ) %>%
  dplyr::rename( result = final ) %>%
  mutate( status = "published" )

analyses_new_processed = 
  analyses_new %>%
  mutate(model = factor(model)) %>%
  select( model, clade, result, inference ) %>%
  mutate( status = "new" )

analyses_processed = 
  rbind( analyses_published_processed, analyses_new_processed ) %>%
  # mutate(software=recode( software, `PhyloBayes MPI`="PhyloBayes") ) %>%
  mutate(model=recode(model,
                    `GTR20`="GTR",
                    `CAT+F81`="Poisson+CAT",
                    `Poisson + CAT`="Poisson+CAT",
                    `GTR + CAT`="GTR+CAT")) %>%
  mutate( 
    model = factor( model, 
      levels=c("WAG", "LG", "GTR", "data partitioning", "Poisson+C60", "WAG+C60", "LG+C60", "Poisson+CAT", "GTR+CAT") )) %>%
  mutate( status = factor(status, levels = c("published", "new")))

p = analyses_processed %>%
	ggplot( aes(x=model, y=clade, col=result, shape=inference) ) +
    facet_wrap( ~ status ) +
		geom_jitter( width = 0.20, height = 0.20, alpha=0.7, size=2 ) +
		scale_colour_manual(values = result_colors) +
		theme_classic() + 
    theme(axis.text.x = element_text(angle=45, hjust=1))+
    xlab("Model") + 
    ylab("Clade")

p

**Fig. 3.**  A summary of phylogenomic results from the previous studies and reanalysis conducted in this study. Each shape represents an phylogenetic analysis that has been analyzed by ML (designated by the circle) or BI inference (designated by the triangle). Analyses with a support of bootstrap values less than 90% or a posterior probability less than 0.95 are considered as unresolved. (A) Analyses transcribed from the literature (related to Supplementary Table XX). (B) New phylogenomic analyses were conducted in this study (related to Supplementary Table XX).

We parsed the `r nrow(analyses_published)` previous phylogenetic analyses from `r length(unique(analyses_published$manuscript))` studies (Fig. 3A and Supplementary Table 2). Among these studies, the main conclusions of five studies strongly favor of Porifera-sister and 10 favor Ctenophora-sister (Extended Data Table 1). Three studies are based entirely on previously published data, and the remainder add data for one or more species.

All published analyses that use models with site-homogeneous equilibrium frequencies (*i.e.*, models that do not include CAT) strongly support Ctenophora-sister, regardless of other factors such as outgroup choice (Fig. 3A). Some analyses that exclude more distant outgroups and use models with site-heterogeneious CAT equilibrium frequencies support Porifera-sister. Because CAT is so parameter rich, many analyses that use a CAT model do not converge, and some of the unconvered analyses have an unresolved animal root. The strong support of Porifera-sister can only be recovered from analyses analyses of protein sequences using Poisson+CAT. Although GTR+CAT model generally has better model adequacy than Poisson+CAT or site-homogeneous models from previous studies[@pisani2015genomic; @feuda2017improved], Porifera-sister is not always supported since several analyses strongly in favor Ctenophora-sister even with only choanofalgellates are used as outgroups (Fig. 3A). Thus, our summary of previous phylogenetic analyses strongly indicated that the trend of the support of Porifera-sister is generally increasing with the use of more complex CAT models and more closely related outgroups in several representative matrices, whereas Ctenophora-sister is supported in most other conditions.

@@

``` {r sensitive, fig.width=7.2, fig.height=7.2}
